---
title: "Lab 1 - Data visualization"
author: "[Kaley Sperling]"
format: pdf
editor: visual
---

## Load Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(viridis)
```

## Exercise 1

```{r}

glimpse(midwest)
```

```{r}
#|histogram-of-population-density-of-counties
ggplot(midwest, aes(x=popdensity)) +
  geom_histogram(binwidth=10000) +
  labs (x="population density",
        y="count",
        title= "population density by county")
```

This is a histogram with a binwidth of 10,000. With this, it is clear to see the data is right-skewed. There appears to be some sort of outlier because there are no counties with a population density of 75000 but there seems to be one with much more than this.

## Exercise 2

```{r}
#| scatterplot-college degree-below-poverty
ggplot(midwest, aes(x=percollege, y=percbelowpoverty, color=state)) +
  geom_point() +
  labs( x="% college educated",
    y="% below poverty line",
    title= "% college educated vs % below poverty line") +
  scale_color_viridis_d(option="D" )
```

## Exercise 3

In my scatterplot, I can see that the majority of counties in midwest states have less than 30% of their population that are college educated and less than 20% below the poverty line. There is one outlier county in WI where very few people are college educated and nearly 50% are below the poverty line. In general, though, Wisconsin has fewer counties with high percentages of people below the poverty line than other states in the Midwest.

## Exercise 4

```{r}

```

## Exercise 5

```{r}

```

## Exercise 6

```{r}

```

## Exercise 7

```{r}

```
